RationalGRL: A Framework for Argumentation and Goal Modeling

### 2018-05-03 Response to reviewers.txt

===== REPLY #2
We agree that many algorithms are similar, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two or more.

===== COMMENT #3
The methodology and the tools looks quite preliminary and without a real validation of the approach. I think this is the weakest part of the paper. Authors should demonstrate the applicability, scalability and usefulness of their approach in the real word. The tool and the modeling language should be used and evaluated by practitioners while they cope with real requirements engineering problems and complexity issues.

===== REPLY #3
We agree with the reviewer that a real validation of the approach is a necessary next step. However, we also would like to point out that an industrial strength case study is a significant amount of work, and would most likely have enough new insights for a separate paper. To give an example, it took Estrada et al. [11] about 9 months to conduct the case study.

Since the main focus of our article is to explain the novel RationalGRL framework in detail, we feel that adding such an extensive study to the current article may be too much content. Still, we agree that having no validation at all is limiting, so we have performed a small-scale user study in order to provide an initial validation of our framework. This empirical study was performed amongst ==> TODO(marcvanzee): ADD NUMBER HERE  <== participants, each with a computer science background.

@@

===== COMMENT #2
The statement on actually using a reduced GRL should be made in the Abstract and Introduction already, not just in a footnote later.

===== REPLY #2
We have included in the abstract and introduction that we use a simplified version. Note that all the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution. This simplification does not in any way influence the case study or any of the other results. Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would not change the essence of our formalism.

===== COMMENT #3
I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL. In general, how does this argumentation approach really differ from one based on beliefs and, in particular, belief revision? Assume that an argument is a belief of some stakeholder revised by another argument, which is a belief of some other stakeholder?

===== REPLY #3
We have added a comparison between arguments in RationalGRL and beliefs in GRL in the related work section, at the beginning of the paragraph "Goal Modelling".

The comparison between argumentation and belief revision is interesting, especially from a theoretical point of view. In the area of knowledge representation, obtaining correspondences between beliefs revision and the dynamics of argumentation theory is a popular topic of research [12,13,14,15]. However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on beliefs in GRL.

@@

===== REPLY #11
We agree with the reviewer that the two mentioned sentences are confusing. We have improved this as follows:
- We have rewritten the first sentence of the abstract as follows: "Goal-oriented requirements modeling approaches aim at capturing intentions of various stakeholders involved in the development of an information system and refining them into low level goals and tasks."
- We have deleted the first sentence of the introduction, since we feel that it is not necessary to explain the definition of Requirements Engineering in the Requirements Engineering Journal.

===== COMMENT #12
Finally, you should run a spell checker, have a careful proof reading of grammar, and make sure that there is no unresolved reference to a figure in a submitted manuscript.

===== REPLY #11
We’ve ran a spell-checker and proof-read the paper carefully. There was one unresolved reference to a figure, which a corrected. Thank you for pointing this out.

######## REFERENCES ########
[1] Y. Elrakaiby, A. Ferrari, P. Spoletini, S. Gnesi, and B. Nuseibeh. Using argumentation to explain ambiguity in requirements elicitation interviews. In Proceedings of the 25th IEEE International Requirements Engineering Conference (RE 2017), pages 51–60. IEEE Press, 2017.
[2] I. Jureta, S. Faulkner, and P. Schobbens. Clear justification of mod- eling decisions for goal-oriented requirements engineering. Re- quirements Engineering, 13(2):87–115, May 2008.
[3] Y. Liu, Y. Su, X. Yin, and G. Mussbacher. Combined goal and feature model reasoning with the user requirements notation and jucmnav. In Requirements Engineering Conference (RE), 2014 IEEE 22nd International, pages 321–322. IEEE, 2014.
[4] P. K. Murukannaiah, A. K. Kalia, P. R. Telangy, and M. P. Singh. Resolving goal conflicts via argumentation-based analysis of com- peting hypotheses. In Proceedings of the 23rd International Re- quirements Engineering Conference (RE 2015), pages 156–165. IEEE, 2015.
[5] M. van Zee, F. Bex, and S. Ghanavati. Rationalization of Goal Models in GRL using Formal Argumentation. In Proceedings of the 23rd International Requirements Engineering Conference (RE’15), pages 220–225. IEEE Press, August 2015
[6] E. S. K. Yu. Towards Modeling and Reasoning Support for Early-Phase Requirements Engineering. In Proceedings of the 3rd IEEE International Symposium on Requirements Engineering, pages 226–235, 1997.

### 2018-05-03 discussion.tex

>>>>

\paragraph{Beliefs in GRL} As mentioned already in the introduction, it is possible in GRL to rationalize models
through so-called \emph{belief} elements. These beliefs are elements that can be attached to an IE with a \emph{belief link}. The content of a belief is a textual description, which can explain, for instance, the rationale for including an IE in the resulting goal model. There are several important differences between GRL beliefs and RationalGRL arguments.

First of all, beliefs in GRL are not used in any of the algorithms to compute the satisfiability of the IEs. This means that a GRL belief can be understood as metadata that can be attached to an element, but it does not have semantics. Argument in RationalGRL, in contrast, play an important role in forming the resulting goal model. If an IE is defeated by an argument, it is not part of the resulting goal model.

Secondly, arguments in RationalGRL can not only provide rationale for IEs in the resulting goal model, but they can also provide rational for those that are rejected, and they can be attached to links as well. This means that Rational arguments are strictly more expressive than GRL beliefs.

Finally, the formation of arguments and counterarguments in RationalGRL is guided by critical questions, which stimulate discussions between stakeholders. This is not possible for GRL beliefs.
